---
title: "Hints"
format:
  html:
    toc: true
---


# Suggested Workflow

1. Look at and take basic notes about final product. WALK.
2. Mock up a grouped bar chart that looks in structure like final, review data structure. DRINK A BEVERAGE.
3. Review inputs. DRAW OR PLAY MUSIC OR GO FOR A BIKE RIDE.
4. Do what you can to get started.  CELEBRATE AND REVIEW.  GO FOR ANOTHER WALK OR SLEEP OR VEG OUT.  HANG WITH THE CAT. 

*I need need need to break this down somehow.  What do I know and currently have onhand ?*

## Look at final product

1. Has 3 BpSs
2. Each BpS has percent of past and present amounts of classes
3. Some formatting


## Mock up of grouped bar chart

* have 3 groups
* have same values per group
* make up data

```{r warnings = FALSE, message = FALSE}
# Load necessary libraries
library(tidyverse)

# Set seed for reproducibility
set.seed(123)

# Create the dataset
data <- data.frame(
  bps = rep(c("ecosystem A", "ecosystem B", "ecosystem C"), each = 10),
  label = rep(c("A", "B", "C", "D", "E", "Agriculture", "Developed", "Water", "UN", "UE"), times = 3),
  ref_cur = rep(c("refPercent", "currentPercent"), each = 15),
  amount = sample(10:200, 30, replace = TRUE)
)

# Plot
fake_plot <-
  ggplot(data, aes(fill=(ref_cur), y=amount, x=label)) +
  geom_col(width = 0.8, position = position_dodge()) +
  coord_flip() +
  facet_grid(. ~bps) +
  labs(
    title = "Succession Classes past and present",
    subtitle = "Fake chart with fake data",
    caption = "Data from landfire.gov",
    x = "",
    y = "Amount (units unknown")

fake_plot

```


* then review data structurem head and tail (in case this clown tried to trick us)
* where do we get this data?


## Review inputs

* What is in them?
* What parts do I need?


### bps_aoi_attributes

```{r}
bps_aoi_attributes <- read_csv("input_data/bps_aoi_attributes.csv")
head(bps_aoi_attributes)
tail(bps_aoi_attributes)
```

* Has total amounts of each BpS
* I probably only need BPS_MODEL, BPS_NAME and some total amount (Freq or ACRES) so I can calculate percents later
* Keep handy just in case
* ALSO, after looking at the combine dataset I might to keep the VALUE field just in case

### bps_model_number_name

```{r}
bps_model_number_name <- read_csv("input_data/bps_model_number_name.csv")
head(bps_model_number_name)
tail(bps_model_number_name)
```

* Not sure why I need this
* This is B.S.

### combine_raw

```{r}
combine_raw <- read_csv("input_data/combine_raw.csv")
head(combine_raw)
tail(combine_raw)
```

* what the?
* I sorted by 'Var 1'.  Wait, I see that Var1 of '11' looks the same as VALUE '11' in the bps_aoi_attributes.  This is getting interesting.  I wonder if Var1 is the same as VALUE in the bps_aoi_attributes.  
* I better keep that VALUE field just in case. 
* **I need to get the current amounts of each class from this somehow and their labels**

### LF16_BPS_200

```{r}
LF16_BPS_200 <- read_csv("input_data/LF16_BPS_200.csv")
head(LF16_BPS_200)
tail(LF16_BPS_200)
```

* This looks like bps_attributes_aoi, but much bigger.  
* I hate this

###  ref_con_modified

```{r}
ref_con_modified <- read_csv("input_data/ref_con_modified.csv")
head(ref_con_modified)
tail(ref_con_modified)
```

* has way more bpss than I need.  Dang will need to filter to my BpSs somehow
* **I will need to pivot longer I think**


### scls_aoi_attributes

```{r}
scls_aoi_attributes <- read_csv("input_data/scls_aoi_attributes.csv")
head(scls_aoi_attributes)
tail(scls_aoi_attributes)
```

* Has labels!  But how to merge?  Lemme look, do the VALUES match up with VAR2 in the combine?  


## Do what you can to get started

1. Pivot the ref_con and try to filter somehow.  I need to freaking filter by the bpss in the area of interest and then just by the top 3.  
**How does that look**
AND AND I need to keep something to merge by later I *think?*  'cause I need to connect the reference and current

2. I think I need to merge in the s-class labels to the combine
3. Oh and merge in thingies from the bps-aoi to the combine
4. Now this is awesome
5. Can I calculate percents of each s-class per BpS?  What the _____ does that give me?  What is this combine thing again?  Better look at the data dictionary


## Get Ref Con Wrangled

```{r}

ref_con <- ref_con_modified %>%
  pivot_longer(!Model_Code, names_to = "refLabel", values_to = "refPercent") %>%
  unite(model_label, c("Model_Code", "refLabel"), remove = FALSE) %>%
  left_join(bps_model_number_name)


# get list of aoi BpS model numbers

aoi_bps_models <- bps_aoi_attributes$BPS_MODEL

#subset ref_con to aoi, keeping all BpSs in the AOI for now
aoi_ref_con <- subset(ref_con, Model_Code %in% aoi_bps_models)


```


## Get Current Wrangled and calculations made

```{r}
#bring in s-class labels
combine <- left_join(combine_raw, 
                  scls_aoi_attributes %>%
                    dplyr::select(2, 4),  
                  by = c("Var2" = "VALUE"))

#bring in bps labels
combine <- left_join(combine, 
                  LF16_BPS_200 %>%
                    dplyr::select(1:4),
                  by = c("Var1" = "VALUE"))

# calculate current sclass percents
combine <- combine %>%
  group_by(Var1, BPS_MODEL) %>%
  mutate(total_count = sum(Freq)) %>%
  mutate(currentPercent = as.integer((Freq/total_count)*100)) %>%
  unite(model_label, c("BPS_MODEL", "LABEL"))

## Note, in general percents for any one BpS should = 100, but may not due to water or 'noise'. They should be close. 

## Also note Jaxson's comment in Slack---there may be missing sclass lables here if a class is not mapped on the current landscape.  For example, old-growth classes are often missing on the current landscape, but were important historically

```


